---
output:
  pdf_document: default
  html_document: default
---

#Final Project

# Question - Can type of Sport be predicted based on a set of variables?

```{r}
require(randomForest)
require(tidyverse)
require(reshape2)
require(scales)
require(ggplot2)
require(moments)
```

```{r}
# Loading dataset
setwd("/Users/rajindersingh/Documents")
```

```{r}
olympics <- read.csv("athlete_events.csv")
```
```{r}
str(olympics)
summary(olympics)
head(olympics)
```

# Omit 'NA'
```{r}
any(is.na(olympics))
olympics <- na.omit(olympics)
any(is.na(olympics))
str(olympics)
summary(olympics)
head(olympics)
```

# Table of categoty proportions for Sex
```{r}
prop.table(table(olympics$Sex))
```

# Number of unique Sport 
```{r}
unique(olympics$Sport)
```
# Comparing Men & Women dataset for Age, Height & Weight
```{r}
boxplot(Age~Sex, data=olympics, col=(c("gold","darkgreen")), 
        main = "Age of Men & Women", xlab="Sex",ylab="Age")


boxplot(Height~Sex, data=olympics, col=(c("gold","darkgreen")), 
        main = "Height of Men & Women", xlab="Sex",ylab="Height")


boxplot(Weight~Sex, data=olympics, col=(c("gold","darkgreen")), 
        main = "Weight of Male & Female", xlab="Sex",ylab="Weight")
```
# The mean of Men’s and Women’s age is almost similar, but the mean for
# Men's height and weight is higher than women.


# Men Athlete_Events - randomForest analysis
```{r}
results <- olympics
str(results)

length(which(results$Sex == "M"))

results.men <- results[results$Sex == "M", ]
str(results.men)

results.men$Sport <- as.factor(results.men$Sport)
str(results.men)
```
```{r}
results.men.rf <- randomForest(results.men[, -13], results.men[, 13])
results.men.rf

importance(results.men.rf)
```

# Remove Event variable since similar to Sport variable
```{r}
results1.men <- results.men[, -14]
str(results1.men)
```

# Athlete_Events - randomForest analysis (Event variable removed)
```{r}
results1.men.rf <- randomForest(results1.men[, -13], results1.men[, 13])
results1.men.rf

importance(results1.men.rf)
```

# Men - histogram with normal curve of age
```{r}
h<-hist(results1.men$Age, breaks=20, col="red", xlab="Age", 
        main="Histogram with Normal Curve of Men's Age") 
xfit<-seq(min(results1.men$Age),max(results1.men$Age),length=40) 
yfit<-dnorm(xfit,mean=mean(results1.men$Age),sd=sd(results1.men$Age)) 
yfit <- yfit*diff(h$mids[1:2])*length(results1.men$Age) 
lines(xfit, yfit, col="blue", lwd=2)
```
# Men - mean and standard deviation of age
```{r}
Ave <- mean(results1.men$Age)
SD <- sd(results1.men$Age)
c(Average = Ave, SD = SD)
```
# Men - normal distribution histogram of age
```{r}
y <-  rnorm(results1.men$Age)
hist(y,main = 'Age Normal Distribution', col = "light green")
```
# calculate standard units
```{r}
z <- scale(results1.men$Age)

# calculate proportion of values within 2 SD of mean
mean(abs(z) < 2)
```

# Men - histogram with normal curve of height
```{r}
h<-hist(results1.men$Height, breaks=20, col="red", xlab="Height", 
        main="Histogram with Normal Curve of Men's Height") 
xfit<-seq(min(results1.men$Height),max(results1.men$Height),length=40) 
yfit<-dnorm(xfit,mean=mean(results1.men$Height),sd=sd(results1.men$Height)) 
yfit <- yfit*diff(h$mids[1:2])*length(results1.men$Height) 
lines(xfit, yfit, col="blue", lwd=2)
```
# Men - mean and standard deviation of height
```{r}
Ave <- mean(results1.men$Height)
SD <- sd(results1.men$Height)
c(Average = Ave, SD = SD)
```
# Men - normal distribution histogram of height
```{r}
y <-  rnorm(results1.men$Height)
hist(y,main = 'Height Normal Distribution', col = "light green")
```
# calculate standard units
```{r}
z <- scale(results1.men$Height)

# calculate proportion of values within 2 SD of mean
mean(abs(z) < 2)
```

# Men - histogram with normal curve of weight
```{r}
h<-hist(results1.men$Weight, breaks=30, col="red", xlab="Weight", 
        main="Histogram with Normal Curve of Men's Weight") 
xfit<-seq(min(results1.men$Weight),max(results1.men$Weight),length=40) 
yfit<-dnorm(xfit,mean=mean(results1.men$Weight),sd=sd(results1.men$Weight)) 
yfit <- yfit*diff(h$mids[1:2])*length(results1.men$Weight) 
lines(xfit, yfit, col="blue", lwd=2)
```
# Men - mean and standard deviation of weight
```{r}
Ave <- mean(results1.men$Weight)
SD <- sd(results1.men$Weight)
c(Average = Ave, SD = SD)
```
# Men - normal distribution histogram of weight
```{r}
y <-  rnorm(results1.men$Weight)
hist(y,main = 'Weight Normal Distribution', col = "light green")
```
# calculate standard units
```{r}
z <- scale(results1.men$Weight)

# calculate proportion of values within 2 SD of mean

mean(abs(z) < 2)
```

# Remove ID, Name, Sex, Team, NOC, Games, Year, Season, City, Medal columns
```{r}
results2.men <- results1.men[, c(-1:-3, -7:-12, -14)]
str(results2.men)
```

# Athlete_Events - randomForest analysis with Age, Height, Weight,and Team
```{r}
results2.men.rf <- randomForest(results2.men[, -4], results2.men[, 4])
options(max.print=999999)
results2.men.rf

importance(results2.men.rf)
```

# randomForest analysis for the following Sport
# (Gymnastics, Athletics, Swimming, Weightlifting, and Equestrianism)
# based on Men's Age, Height, Weight, and Team
```{r}  
length(which(olympics$Sport == "Gymnastics"))
olympics.gymnastics <- olympics[olympics$Sport == "Gymnastics", ]
str(olympics.gymnastics)

length(which(olympics$Sport == "Athletics"))
olympics.athletics <- olympics[olympics$Sport == "Athletics", ]
str(olympics.athletics)

length(which(olympics$Sport == "Swimming"))
olympics.swimming <- olympics[olympics$Sport == "Swimming", ]
str(olympics.swimming)

length(which(olympics$Sport == "Weightlifting"))
olympics.weightlifting <- olympics[olympics$Sport == "Weightlifting", ]
str(olympics.weightlifting)

length(which(olympics$Sport == "Equestrianism"))
olympics.equestrian <- olympics[olympics$Sport == "Equestrianism", ]
str(olympics.equestrian)

olympics.combine <- rbind(olympics.gymnastics, olympics.swimming, olympics.weightlifting, 
                          olympics.athletics, olympics.equestrian)
str(olympics.combine)
```
```{r}
olympics.combine1 <- olympics.combine[, c(-1:-3, -7:-12, -14, -15)]
str(olympics.combine1)
```
```{r}
olympics.combine1$Sport <- as.factor(olympics.combine1$Sport)
str(olympics.combine1)
```
```{r}
olympics.combine1.rf <- randomForest(olympics.combine1[,-4], olympics.combine1[,4])
olympics.combine1.rf

importance(olympics.combine1.rf)
```

# Predicting Sport based on randomForest analysis
# Row 1000 & 4000
```{r}
olympics.combine1[1000,]
olympics.combine1[4000,]
```

#Predicting Row 1000 & 4000
```{r}
predict(olympics.combine1.rf, olympics.combine1[1000,-5])
predict(olympics.combine1.rf, olympics.combine1[4000,-5])
```

# Perdicting for Age = 30, Height = 170, Weight = 50
```{r}
predict(olympics.combine1.rf, c(30, 170, 50))
```

# Heatmap with relative change equally across all variables
# Rescale Age, Height and Weight 0 to 100
```{r}
str(olympics.combine1)
olympics.combine1$Age <- rescale(olympics.combine1$Age, to = c(0, 100))
olympics.combine1$Height <- rescale(olympics.combine1$Height, to = c(0, 100))
olympics.combine1$Weight <- rescale(olympics.combine1$Weight, to = c(0, 100))
str(olympics.combine1)
```

# Melt Age, Weight, Height against Sport
```{r}
melt <- reshape::melt(olympics.combine1, id=c("Sport"))
str(melt)

gplot <- ggplot(melt, aes(x = Sport, y = variable,
                             group = variable, fill = value))
gplot <- gplot + geom_tile()
gplot <- gplot + scale_fill_gradient(low = 'blue', high = 'red')
gplot <- gplot + ggtitle("Heatmap of Men's Age, Height Weight and Sport",
                         subtitle = "Relative change equally across Age, Height, and Weight")
gplot <- gplot + labs(x = "Sport", y='Variables') 
gplot
```

# Scatter plot Men's Age, Height, Weight, 
# and Sport (Athletics, Equestrianism, Gymnastics, Swimming, & Weightlifting)
```{r}
str(olympics.combine)
```
```{r}
olympics.combine2 <- olympics.combine[, c(-1:-3, -7:-12, -14, -15)]
str(olympics.combine2)
```

```{r}
gplot <- ggplot(olympics.combine2, aes(x=Height, y=Weight))
gplot <- gplot + geom_point(aes(size=Age, color=as.factor(Sport)))
gplot <- gplot + stat_smooth(method = "lm",  aes(x=Height, y=Weight))
gplot <- gplot + labs(x = 'Height', y = 'Weight', size = 'Age',
                      color = "Sport") 
gplot <- gplot + ggtitle("Scatter plot of Men's Height, Weight, Age & Sport")
gplot
```

# randomForest analysis for the following Sport
# (Equestrianism, Gymnastics, Swimming, and Weightlifting)
# based on Men's Age, Height, Weight, and Team
```{r}
olympics.combine3 <- rbind(olympics.gymnastics, olympics.swimming, olympics.weightlifting, olympics.equestrian)
str(olympics.combine3)
```
```{r}
olympics.combine4 <- olympics.combine3[, c(-1:-3, -7:-12, -14, -15)]
str(olympics.combine4)
```
```{r}
olympics.combine4$Sport <- as.factor(olympics.combine4$Sport)
str(olympics.combine4)
```
```{r}
olympics.combine4.rf <- randomForest(olympics.combine4[,-4], olympics.combine4[,4])
olympics.combine4.rf

importance(olympics.combine4.rf)
```

# Scatter plot Men's Age, Height, Weight, 
# and Sport (Equestrianism, Gymnastics, Swimming, & Weightlifting)
```{r}
str(olympics.combine4)
```
```{r}
gplot <- ggplot(olympics.combine4, aes(x=Height, y=Weight))
gplot <- gplot + geom_point(aes(size=Age, color=as.factor(Sport)))
gplot <- gplot + stat_smooth(method = "lm",  aes(x=Height, y=Weight))
gplot <- gplot + labs(x = 'Height', y = 'Weight', size = 'Age',
                      color = "Sport") 
gplot <- gplot + ggtitle("Scatter plot of Men's Height, Weight, Age & Sport")
gplot
```

# Create train and test data sets
# based on train dataset for the following Sport
# (Equestrianism, Gymnastics, Swimming, and Weightlifting)
# based on Men's Age, Height, Weight, and Team
```{r}
str(olympics.combine4)
```
```{r}
split.olympics.combine4 <- sample(nrow(olympics.combine4), nrow(olympics.combine4) * .67)
olympics.combine4.Train <- olympics.combine4[split.olympics.combine4, ]
olympics.combine4.Test <- olympics.combine4[-split.olympics.combine4, ]
str(olympics.combine4.Train)
str(olympics.combine4.Test)
```

# randomForest analysis based on train dataset for the following Sport
# (Equestrianism, Gymnastics, Swimming, and Weightlifting)
# based on Men's Age, Height, Weight, and Team
```{r}
olympics.combine4.Train$Sport <- as.factor(olympics.combine4.Train$Sport)
str(olympics.combine4.Train)
```
```{r}
olympics.combine4.Test$Sport <- as.factor(olympics.combine4.Test$Sport)
str(olympics.combine4.Test)
```
```{r}
olympics.combine4.Train.rf <- randomForest(olympics.combine4.Train[,-4], olympics.combine4.Train[,4])
olympics.combine4.Train.rf

importance(olympics.combine4.Train.rf)
```

# Predicting Sport based on randomForest analysis

# Row 500 & 1300
```{r}
olympics.combine4.Test[500,]
olympics.combine4.Test[1300,]
```

#Predicting Row 5000 & 1300
```{r}
predict(olympics.combine4.Train.rf, olympics.combine4.Test[500,-4])
predict(olympics.combine4.Train.rf, olympics.combine4.Test[1300,-4])
```

# Perdicting for Age = 30, Height = 170, Weight = 50
```{r}
predict(olympics.combine4.Train.rf, c(30, 170, 50))
```

# Plot on random Forest model
```{r}
plot(olympics.combine4.Train.rf)
```

# lm() regression models
```{r}
str(olympics.combine4)
olympics.combine4$Sport <- as.factor(olympics.combine4$Sport)
str(olympics.combine4)
```
```{r}
olympics.combine4$Sport1 <- c(0:0)
str(olympics.combine4)
```
```{r}
df.sport <- c('Equestrianism', 'Gymnastics', 'Swimming', 'Weightlifting')
str(df.sport)
```
```{r}
olympics.combine4$Sport1 <- match(olympics.combine4$Sport, df.sport, nomatch = 0)
```
```{r}
knitr::kable(head(olympics.combine4[which(olympics.combine4$Sport == "Equestrianism"),]))

knitr::kable(head(olympics.combine4[which(olympics.combine4$Sport == "Gymnastics"),]))

knitr::kable(head(olympics.combine4[which(olympics.combine4$Sport == "Swimming"),]))

knitr::kable(head(olympics.combine4[which(olympics.combine4$Sport == "Weightlifting"),]))
```
```{r}
predict.sport.age <- lm(Sport1 ~ Age, olympics.combine4)
summary(predict.sport.age)
```
```{r}
predict.sport.height <- lm(Sport1 ~ Height, olympics.combine4)
summary(predict.sport.height)
```
```{r}
predict.sport.weight <- lm(Sport1 ~ Weight, olympics.combine4)
summary(predict.sport.weight)
```
```{r}
predict.sport.height.weight <- lm(Sport1 ~ Height + Weight, olympics.combine4)
summary(predict.sport.height.weight)
```
```{r}
predict.sport <- lm(Sport1 ~ Age + Height + Weight, olympics.combine4)
summary(predict.sport)
```

